Document Indexing Using Aho-Corasick State Machines

### 2012-02-17 report/article.tex

%----------------------------------------
% Design
%----------------------------------------
\section{Implementation}
\label{sec:implementation}
The Aho-Corasick indexer was implemented using Java and MySQL
relational database. Upon startup, the indexer reads a list of index
keywords and synonyms from the database and constructs an Aho-Corasick
state machine. The indexer then enters a loop in which it uses the
state machine to scan batches of documents retrieved from the
database. The document batch size is varied manually to match the
physical capabilites of the host. A possible enhancement to our
implementation could include an adaptive loading component. The output
of the state machine is used to augment a global index maintained in
the database.

@@

%----------------------------------------
% Comparison with Naive Index Builder
%----------------------------------------
\subsection{Comparison with Na\"{i}ve Index Builder}
\label{sec:naive}
In order to quantify the benefits of using the Aho-Corasick algorithm
for indexing, our indexer performace was compared to that of a
na\"{i}ve implementation. The na\"{i}ve indexer builder has the same
input as our system i.e. a list of keywords and a list of documents,
and the same output i.e. an inverted index corresponding to the
keywords and documents. Algorithm \ref{alg:invertedindex} is used in
principal for both implementations. However, whereas the Aho-Corasick
implementation uses a state machine to scan, the na\"{\i}ve
indexer tokenizes the document and compares each token to the keyword
dictionary.

@@

Algorithm \ref{alg:invertedindex} has a complexity proportional to the
number of keywords $n$, the number of document $m$, and the average
document size $\bar{c}$. For our chosen architecture, execution time
is also effected by database latency. In the case of the
na\"{\i}ve indexer, we perform one database transaction per keyword
and in the Aho-Corasick implementation, we perform a database
transaction per document. For analysis purposes, we designate the
average database latency per transaction as $\bar{t}$.

@@

To examine the effect of document size on idexer performance, the number of keywords was
fixed at 1,000, the number of documents at 5,000, and tests run
against corpora composed of varying average length. From equations
\ref{eqn:timecomplexworstcase} and \ref{eqn:timecomplexsm}, we would
expect both implementations to behave linearly with repect to average
documetn size and to differ by constant scaling factor $n$, the size
of the keyword dictionary. The experiment results are shown in figure
\ref{fig:naivedocumentsize} and support this analysis. Also, we can observe
that for smaller documents, the na\"{\i}ve method
performs better. This is attributable to the fact that for smaller
documents, the database advantage of the na\"{\i}ve implementation is
enough to overcome the keyword dictionary size scaling factor. To see
this effect more clearly, experiment results are plotted on a
logarithmic scale in figure \ref{fig:naivedocumentsizelog}.

\begin{figure}[ht]
  \begin{minipage}[b]{0.5\linewidth}
    \centering
    \includegraphics[width=\textwidth]{naivedocumentsize}
    \caption{Time to build an inverted index as a function of the
      average document size. Time taken is shown in miliseconds on the
      y-axis and the average document size in characters on the x-axis}
    \label{fig:naivedocumentsize}
  \end{minipage}
  \hspace{0.5cm}
  \begin{minipage}[b]{0.5\linewidth}
    \centering
    \includegraphics[width=\textwidth]{naivedocumentsizelog}
    \caption{Time to build an inverted index as a function of the
      average document size. Time taken is shown in milliseconds on the
      y-axis and the average document size in characters on the
      logarithmic x-axis}
    \label{fig:naivedocumentsizelog}
  \end{minipage}
\end{figure}

xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx

However, we would have expected the state machine method not to depend
on the number of keyword. The reason why it’s not the case is because
when we don’t have any keywords in a document, we don’t access the
database. Knowing that the database access is very time consuming,
increasing the number of keyword increases the probability of finding a
keyword into each document and so increase the time complexity.

@@

As expected, the time complexity is linearly related to the number of
documents being indexed. The experiment results, in this respect, are
shown in figure \ref{fig:naivesizecorpus} and on a logarithmic scale
in figure \ref{fig:naivesizecorpuslog}. For the na\"{i}ve
implementation:

\[Time \propto (docNum * maxLenght + conn) * keyNum\]

and in our Aho-Corasick state machine method:

\[Time \propto docNum * maxLenght + conn * docNum\]

@@

%----------------------------------------
% Parallelization
%----------------------------------------
\subsection{Parallelization}
\label{parallelization}
Given the modular architecture of our indexer, it is possible to
deploy multiple instances of the program to work concurrently on the same
indexing task. The instances interact via the database which
synchronizes their operations on the corpus and the index itself. To
investigate the effect of parallelization on the indexer operation, the
indexer was deployed using a single database server storing documents,
keywords, and the resulting index; four indexer nodes each running a
single instance of the indexer program.

In this experiment a corpus of 100,000 documents and 1000 keywords were
used, each document 5000 words in length. Tests were run using various
node combinations and the time taken to index the corpus was measured and
normalized to a throughput metric (documents/second) for convenient
comparison. Test results using the same number of nodes were averaged
to account for differences in the physical capabilities and
configuration of the worker nodes. Next, a series of tests was
performed using a corpus of 800,000 documents to investigate the
effect of corpus size on performance. The results of this experiment
are shown in figure \ref{fig:parallelization}.

>>>>

%----------------------------------------
% Design
%----------------------------------------
\section{Implementation}
\label{sec:implementation}
The Aho-Corasick indexer was implemented using Java and MySQL
relational database. Upon startup, the indexer reads a list of index
keywords and synonyms from the database and constructs an Aho-Corasick
state machine. The indexer then enters a loop in which it uses the
state machine to scan batches of documents retrieved from the
database. The document batch size is varied manually to match the
physical capabilities of the host. A possible enhancement to our
implementation could include an adaptive loading component. The output
of the state machine is used to augment a global index maintained in
the database.

@@

%----------------------------------------
% Comparison with Naive Indexer
%----------------------------------------
\subsection{Comparison with Na\"{i}ve Indexer}
\label{sec:naive}
In order to quantify the benefits of using the Aho-Corasick algorithm
for indexing, our indexer performance was compared to that of a
na\"{i}ve implementation. The na\"{i}ve indexer builder has the same
input as our system i.e. a list of keywords and a list of documents,
and the same output i.e. an inverted index corresponding to the
keywords and documents. Algorithm \ref{alg:invertedindex} is used in
principal for both implementations. However, whereas the Aho-Corasick
implementation uses a state machine to scan, the na\"{\i}ve
indexer tokenizes the document and compares each token to the keyword
dictionary.

@@

Algorithm \ref{alg:invertedindex} has a complexity proportional to the
number of keywords $n$, the number of documents $m$, and the average
document size $\bar{c}$. For our chosen architecture, execution time
is also effected by database latency. In the case of the
na\"{\i}ve indexer, we perform one database transaction per keyword
and in the Aho-Corasick implementation, we perform a database
transaction per document. For analysis purposes, we designate the
average database latency per transaction as $\bar{t}$.

@@

In the context of these equations, we describe and interpret our
empirical results in the following subsections, testing for the effects
of document size $c$, number of keywords $n$, and number of documents
$m$.

%----------------------------------------
% Document Size
%----------------------------------------
\subsubsection{Document Size}
To examine the effect of document size on indexer performance, the
number of keywords was fixed at 1,000, the number of documents at
5,000, and tests run against corpora composed of varying average
length. From equations \ref{eqn:timecomplexworstcase} and
\ref{eqn:timecomplexsm}, we would expect both implementations to
behave linearly with respect to average document size and to differ by
a constant scaling factor $n$, the size of the keyword dictionary. The
experiment results are shown in figure \ref{fig:naivedocumentsize} and
support this analysis. Also, we observe that for smaller
documents, the na\"{\i}ve method performs better. This is attributable
to the fact that for smaller documents, the database advantage of the
na\"{\i}ve implementation is enough to overcome the keyword dictionary
size scaling factor. To see this effect more clearly, experiment
results are plotted on a logarithmic scale in figure
\ref{fig:naivedocumentsizelog}.

\begin{figure}[ht]
  \begin{minipage}[b]{0.5\linewidth}
    \centering
    \includegraphics[width=\textwidth]{naivedocumentsize}
    \caption{Time to build an inverted index as a function of the
      average document size. Time taken is shown in milliseconds on the
      y-axis and the average document size in characters on the x-axis}
    \label{fig:naivedocumentsize}
  \end{minipage}
  \hspace{0.5cm}
  \begin{minipage}[b]{0.5\linewidth}
    \centering
    \includegraphics[width=\textwidth]{naivedocumentsizelog}
    \caption{Time to build an inverted index as a function of the
      average document size. Time taken is shown in milliseconds on the
      y-axis and the average document size in characters on the
      logarithmic x-axis}
    \label{fig:naivedocumentsizelog}
  \end{minipage}
\end{figure}

%----------------------------------------
% Keywords
%----------------------------------------
\subsubsection{Number of Keywords}
To examine the effect of keyword dictionary size on indexer
performance, the number of documents was fixed at 5,000, and tests run
using varying numbers of keywords. From equation
\ref{eqn:timecomplexworstcase}, we would expect the na\"{\i}ve
implementation to have a linear dependency on the number of keywords
and this is supported by the test results. From equation
\ref{eqn:timecomplexsm} we would not expect the Aho-Corasick
implementation to depend on the number of keyword however, the test
results seem to indicate a linear dependency. This can be attributed
to the fact that the larger the keyword dictionary, the more likely we
are to find matches in a document and incur the expensive database
latency cost.

@@

%----------------------------------------
% Corpus Size
%----------------------------------------
\subsubsection{Number of Documents}
From equations \ref{eqn:timecomplexworstcase} and
\ref{eqn:timecomplexsm}, the time complexity is linearly related to
the number of documents being indexed. The experiment results, in this
respect, are shown in figure \ref{fig:naivesizecorpus} and on a
logarithmic scale in figure \ref{fig:naivesizecorpuslog}.

@@

%----------------------------------------
% Parallelization
%----------------------------------------
\subsection{Parallelization}
\label{parallelization}
Given the modular architecture of our indexer, it is possible to
deploy multiple instances of the program to work concurrently on the same
indexing task. The instances interact via the database which
synchronizes their operations on the corpus and the index itself. To
investigate the effect of parallelization on the indexer operation, the
indexer was deployed using a single database server storing documents,
keywords, and the resulting index. Four indexer nodes were used, each
running a single instance of the indexer program.

In this experiment a corpus of 100,000 documents and 1000 keywords were
used, each document 5000 words in length. Tests were run using various
node combinations and the time taken to index the corpus was measured and
normalized to a throughput metric (documents/second) for convenient
comparison and to account for differences in the physical capabilities
and configuration of the worker nodes. A series of tests were
performed using a corpus of 800,000 documents to investigate the
effect of corpus size on performance. The results of this experiment
are shown in figure \ref{fig:parallelization}.
